Point estimates comparison for the simplified Naomi model

### 2023-12-21 src/naomi-simple_point-estimates/point-estimates.Rmd
Update point estimate plots for thesis. Just need to add LEB into these

```{r}
df_tmb <- lapply(tmb$fit$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "TMB")

df_aghq <- lapply(aghq$quad$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "PCA-AGHQ")

@@

```{r}
df_plot <- df %>%
  pivot_longer(cols = c("sd", "mean"), names_to = "indicator", values_to = "estimate") %>%
  pivot_wider(values_from = "estimate", names_from = "method") %>%
  pivot_longer(cols = c("TMB", "PCA-AGHQ"), names_to = "method", values_to = "approximate") %>%
  rename("truth" = "NUTS") %>%
  mutate(
    indicator = fct_recode(indicator, "Posterior mean estimate" = "mean", "Posterior SD estimate" = "sd"),
    method = fct_relevel(method, "TMB", "PCA-AGHQ")
  )

@@

Add difference in metric between TMB and PCA-AGHQ:

@@

#' Add as a naming column for plot
df_metrics <- df_metrics %>%
  left_join(
    df_metrics_pct,
    by = c("indicator", "type")
  ) %>%
  mutate(
    label = ifelse(
      method == "PCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  )

@@

df_plot_latent %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_point(shape = 1, alpha = 0.4) +
  facet_grid(indicator ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
  geom_text(data = df_metrics_latent, aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5) +
  labs(x = "NUTS", y = "Approximation - NUTS") +
  theme_minimal()

@@

mean_plot_latent <- df_plot_latent %>%
  filter(indicator == "Posterior mean estimate") %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_jitter(shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  lims(y = c(-0.4, 0.4)) +
  facet_grid(indicator ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
  geom_text(
    data = filter(df_metrics_latent, indicator == "Posterior mean estimate"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "NUTS", y = "Approximation - NUTS") +
  theme_minimal()

@@

sd_plot_latent <- df_plot_latent %>%
  filter(indicator == "Posterior SD estimate") %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_jitter(shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  lims(y = c(-0.6, 0.6)) +
  facet_grid(indicator ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
  geom_text(
    data = filter(df_metrics_latent, indicator == "Posterior SD estimate"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "NUTS", y = "Approximation - NUTS") +
  theme_minimal()

@@

```{r}
df_plot_latent %>%
  filter(indicator == "Posterior mean estimate") %>%
  mutate(diff = abs(approximate - truth)) %>%
  arrange(desc(diff)) %>%
  select(indicator, par, id, method, truth, approximate, diff) %>%
  DT::datatable()
```

@@

```{r}
df_plot_latent %>%
  filter(indicator == "Posterior SD estimate") %>%
  mutate(diff = abs(approximate - truth)) %>%
  arrange(desc(diff)) %>%
  select(indicator, par, id, method, truth, approximate, diff) %>%
  DT::datatable()
```

@@

```{r}
df_plot_output <- filter(df_plot, type == "output")
df_metrics_output <- filter(df_metrics, type == "output")
cbpalette <- c("#56B4E9","#009E73", "#E69F00", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

@@

df_plot_output %>%
  ggplot() +
  geom_point(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4) +
  scale_color_manual(values = cbpalette) +
  facet_grid(indicator ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
  geom_text(data = df_metrics_output, aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5) +
  labs(x = "NUTS", y = "Approximation - NUTS", col = "Output") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

mean_plot_output <- df_plot_output %>%
  filter(indicator == "Posterior mean estimate") %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = cbpalette) +
  facet_grid(indicator ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
  geom_text(
    data = filter(df_metrics_output, indicator == "Posterior mean estimate"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "NUTS", y = "Approximation - NUTS", col = "Output") +
  theme_minimal()

@@

sd_plot_output <- df_plot_output %>%
  filter(indicator == "Posterior SD estimate") %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = cbpalette) +
  facet_grid(indicator ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
  geom_text(
    data = filter(df_metrics_output, indicator == "Posterior SD estimate"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "NUTS", y = "Approximation - NUTS", col = "Output") +
  theme_minimal()

@@

df_metrics_output_split <- df_metrics_output_split %>%
  left_join(
    df_metrics_output_split_pct,
    by = c("indicator", "type", "par")
  ) %>%
  mutate(
    label = ifelse(
      method == "PCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  )

plot_outputs <- function(.indicator, .par, col) {
  df_plot_output %>%
  filter(indicator == .indicator, par == .par) %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c(col, "white", "white")) +
  facet_grid(indicator + par ~ method) +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
  geom_text(
    data = filter(df_metrics_output_split, indicator == .indicator, par == .par),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "NUTS", y = "Approximation - NUTS", col = "Output") +
  theme_minimal() +
  theme(legend.position = "none")
}

mean_prev_plot <- plot_outputs(.indicator = "Posterior mean estimate", .par = "HIV prevalence", col = "#E69F00")
mean_inc_plot <- plot_outputs(.indicator = "Posterior mean estimate", .par = "HIV incidence", col = "#009E73")
mean_art_plot <- plot_outputs(.indicator = "Posterior mean estimate", .par = "ART coverage", col = "#56B4E9")
sd_prev_plot <- plot_outputs(.indicator = "Posterior SD estimate", .par = "HIV prevalence", col = "#E69F00")
sd_inc_plot <- plot_outputs(.indicator = "Posterior SD estimate", .par = "HIV incidence", col = "#009E73")
sd_art_plot <- plot_outputs(.indicator = "Posterior SD estimate", .par = "ART coverage", col = "#56B4E9")

#' Put together
combine_three_plots <- function(plot1, plot2, plot3) {
  plot1alt <- plot1 +
    theme(legend.position = "none")

@@

  three_plots <- cowplot::plot_grid(plot1alt, plot2alt, plot3alt, ncol = 1)
  cowplot::plot_grid(y_axis, three_plots, ncol = 2, rel_widths = c(1, 35))
}

combine_three_plots(mean_prev_plot, mean_inc_plot, mean_art_plot)
ggsave("mean-alt-output.png", h = 7, w = 6.25)

combine_three_plots(sd_prev_plot, sd_inc_plot, sd_art_plot)
ggsave("sd-alt-output.png", h = 7, w = 6.25)
```

>>>>

```{r}
df_tmb <- lapply(tmb$fit$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "GEB")

df_aghq <- lapply(aghq$quad$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "GPCA-AGHQ")

@@

```{r}
df_plot <- df %>%
  pivot_longer(cols = c("sd", "mean"), names_to = "indicator", values_to = "estimate") %>%
  pivot_wider(values_from = "estimate", names_from = "method") %>%
  pivot_longer(cols = c("GEB", "GPCA-AGHQ"), names_to = "method", values_to = "approximate") %>%
  rename("truth" = "NUTS") %>%
  mutate(
    indicator = fct_recode(indicator, "Posterior mean" = "mean", "Posterior SD" = "sd"),
    method = fct_relevel(method, "GEB", "GPCA-AGHQ")
  )

@@

Add difference in metric between GEB and GPCA-AGHQ:

@@

#' Add as a naming column for plot
df_metrics <- df_metrics %>%
  left_join(
    df_metrics_pct,
    by = c("indicator", "type")
  ) %>%
  mutate(
    label = ifelse(
      method == "GPCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  )

@@

df_plot_latent %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_point(shape = 1, alpha = 0.4) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(data = df_metrics_latent, aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5) +
  labs(x = "Estimate", y = "Absolute difference to NUTS") +
  theme_minimal()

@@

mean_plot_latent <- df_plot_latent %>%
  filter(indicator == "Posterior mean") %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_jitter(shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  lims(y = c(-0.4, 0.4)) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_latent, indicator == "Posterior mean"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference to NUTS") +
  theme_minimal()

@@

sd_plot_latent <- df_plot_latent %>%
  filter(indicator == "Posterior SD") %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_jitter(shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  lims(y = c(-0.6, 0.6)) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_latent, indicator == "Posterior SD"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference to NUTS") +
  theme_minimal()

@@

```{r}
df_plot_latent %>%
  filter(indicator == "Posterior mean") %>%
  mutate(diff = abs(approximate - truth)) %>%
  arrange(desc(diff)) %>%
  select(indicator, par, id, method, truth, approximate, diff) %>%
  DT::datatable()
```

@@

```{r}
df_plot_latent %>%
  filter(indicator == "Posterior SD") %>%
  mutate(diff = abs(approximate - truth)) %>%
  arrange(desc(diff)) %>%
  select(indicator, par, id, method, truth, approximate, diff) %>%
  DT::datatable()
```

@@

```{r}
df_plot_output <- filter(df_plot, type == "output")
df_metrics_output <- filter(df_metrics, type == "output")

@@

df_plot_output %>%
  ggplot() +
  geom_point(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4) +
  scale_color_manual(values = c("#56B4E9","#009E73", "#CC79A7")) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(data = df_metrics_output, aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5) +
  labs(x = "Estimate", y = "Absolute difference to NUTS", col = "Output") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

mean_plot_output <- df_plot_output %>%
  filter(indicator == "Posterior mean") %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c("#56B4E9","#009E73", "#CC79A7")) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_output, indicator == "Posterior mean"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference to NUTS", col = "Output") +
  theme_minimal()

@@

sd_plot_output <- df_plot_output %>%
  filter(indicator == "Posterior SD") %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c("#56B4E9","#009E73", "#CC79A7")) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_output, indicator == "Posterior SD"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference to NUTS", col = "Output") +
  theme_minimal()

@@

df_metrics_output_split <- df_metrics_output_split %>%
  left_join(
    df_metrics_output_split_pct,
    by = c("indicator", "type", "par")
  ) %>%
  mutate(
    label = ifelse(
      method == "GPCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  )

plot_outputs <- function(.indicator, .par, col) {
  df_plot_output %>%
  filter(indicator == .indicator, par == .par) %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c(col, "white", "white")) +
  facet_grid(indicator + par ~ method, labeller = labeller(indicator = function(x) {rep("", length(x))})) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4)) +
  geom_text(
    data = filter(df_metrics_output_split, indicator == .indicator, par == .par),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference to NUTS", col = "Output") +
  theme_minimal() +
  theme(legend.position = "none", panel.spacing.x = unit(6, "mm"))
}

mean_prev_plot <- plot_outputs(.indicator = "Posterior mean", .par = "HIV prevalence", col = "#CC79A7")
mean_inc_plot <- plot_outputs(.indicator = "Posterior mean", .par = "HIV incidence", col = "#009E73")
mean_art_plot <- plot_outputs(.indicator = "Posterior mean", .par = "ART coverage", col = "#56B4E9")
sd_prev_plot <- plot_outputs(.indicator = "Posterior SD", .par = "HIV prevalence", col = "#CC79A7")
sd_inc_plot <- plot_outputs(.indicator = "Posterior SD", .par = "HIV incidence", col = "#009E73")
sd_art_plot <- plot_outputs(.indicator = "Posterior SD", .par = "ART coverage", col = "#56B4E9")

combine_three_plots <- function(plot1, plot2, plot3, subtitle) {

  plot1alt <- plot1 + theme(legend.position = "none") + labs(subtitle = subtitle)

  plot2alt <- plot2 +
    theme(
      strip.background = element_blank(),
      strip.text.x = element_blank(),

      legend.position = "none"
    )

@@

  three_plots <- cowplot::plot_grid(plot1alt, plot2alt, plot3alt, ncol = 1, rel_heights = c(1.25, 1, 1))
  cowplot::plot_grid(y_axis, three_plots, ncol = 2, rel_widths = c(1, 35))
}

combine_three_plots(mean_prev_plot, mean_inc_plot, mean_art_plot, subtitle = "Posterior mean")
ggsave("mean-alt-output.png", h = 7, w = 6.25)

combine_three_plots(sd_prev_plot, sd_inc_plot, sd_art_plot, subtitle = "Posterior SD")
ggsave("sd-alt-output.png", h = 7, w = 6.25)
```
